---
title: "midterm_key_oct_8"
author: "Rob Wells"
date: "2025-10-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# JOUR405: Statistics for Journalists
## Midterm Exam - Fall 2025

###Luke Rowe

For this exam, you'll analyze several datasets using R and the statistical concepts we've covered in class. Load the tidyverse before beginning, then complete each task. Write your code in the provided blocks and answer the questions in complete sentences.

# Load tidyverse
```{r}
library(tidyverse)
```

## Part 1: Restaurant Health Inspections (25 points)

The first dataset contains restaurant health inspection scores from Montgomery County. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv`

The dataset includes the name of the establishment, the number of points for critical and non-critical areas, the total points, maximum points possible and the compliance score and grade.

## Task 1: Read the data**
```{r}
#your code here 
restaurant_inspections <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv")
```


## Task 2: Calculate the mean and standard deviation maximum and minimum of compliance scores. Put the results in a new dataframe called "compliance_stats." 
```{r}
#your code here
compliance_stats <- mutate(restaurant_inspections, mean = mean(compliance_score), sd = sd(compliance_score), minimum = min(compliance_score), maximum = max(compliance_score))
```

## Task 3: Create a histogram of the compliance scores with a vertical line showing the mean. Include a headline. Include a headline and label the x and y axes.
```{r}
#your code here
compliance_stats |> 
  ggplot(aes(compliance_score)) +
  geom_histogram() +
  geom_vline(aes(xintercept = mean(compliance_score)), color = "red", linetype = "dashed", size = 1) +
  labs(title = "Compliance Scores", x = "Score", y = "Frequency")


```

## Task 4: Write 2-3 sentences interpreting what the standard deviation and histogram tell us about the distribution of compliance scores. What would be newsworthy about this distribution? What's the story here? 

**Write your answer here:**
The standard deviation tells us that that 2/3 of the data lies between 5.826 less than and greater than the mean of 96.30665. This means that a majority of the compliance scores are above 90, and the historgram reinforces this by visually showing where a majority of the data lies. I think that this is newsworthy, and a story could highlight the high level of compliance from restaurants in the area, and perhaps even rank the top ten.

## Part 2: High School Athletics (30 points)

The second dataset shows participation numbers in high school sports across Maryland counties in 2024, broken down by sport and gender. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_hs_participation.csv`

# Task 5: Load the high school participation data
```{r}
#your code here
high_school_sports <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_hs_participation.csv")
```


#Task 6: Add two columns called total and girls_pct.  
Add two columns called total and girls_pct using mutate(), with the total adding together boys and girls and girls_pct being the percentage of the total represented by girls participants. 
```{r}
#your code here
mutate(high_school_sports, total = rowSums(across(3:4)), girls_pct = girls/total * 100)
```

#Task 7: Calculate correlation between boys and girls participation
```{r}
#your code here
high_school_sports |> summarize(correlation = cor(boys, girls, method = "pearson"))
```

#Task 8: Create a scatterplot showing this relationship, adding a line of best fit. Include a headline and label the x and y axes.
```{r}
#your code here
high_school_sports |>
  ggplot() +
  geom_point(aes(x=boys, y=girls)) +
  geom_smooth(aes(x=boys, y=girls), method = "lm") +
  labs(title = "High School Sports Participation", x = "Boys", y = "Girls")

```


# Task 9: In 2-3 sentences, explain what the correlation coefficient and scatterplot reveal about gender equity in Maryland high school sports participation. How do you interpret the school districts that are below the line vs those that are above? Which school districts are most worth examining further, and why? (10 points)

**Write your answer here:**
The correlation coefficent and scatter plot both show a strong correlation between boys' and girls' participation in high school sports. The coefficient is .98, extremely high when considering that .50 is considered notable. School districts that are below the line have higher boys' participation than girls' participation, and these district should be examined further to reveal why this discrepency has occurred.

## Part 3: Maryland Car Theft Rates (25 points)

Your editor has assigned you a story about car thefts in Maryland and wants you to analyze the data to find out which counties have the highest rates. The fourth dataset contains car theft statistics for Maryland counties in 2023 and population. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv` and complete the following tasks:

#Task 10: Load the car thefts data
```{r}
#your code here
car_thefts <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv")
```

#Task 11: Using mutate, add a column that calculates the rate of car thefts for each county - you need to choose the per capita rate 
```{r}
#your code here
car_thefts2 <- mutate(car_thefts, rate_of_theft_2023 = 2023/population * 100)
```

#Task 12: Calculate the median car theft rate and the total number of car thefts statewide. Which counties have rates above the median? 
```{r}
#your code here
car_thefts2 |> summarize(median_rate_2023 = median(rate_of_theft_2023))
```


#Task 13: Write 2-3 sentences describing what these calculations reveal about the distribution of car thefts in Maryland. What's the lede of a story about your findings? 

**Write your answer here:**
The median rate being 1.8 tells me that for every 100 people, almost 2 will experience a car theft in the state of Maryland. This can serve as a strong lede, as long as it is provided with contextual data to back up how this number was found.

## Part 4: Data Analysis Scenario (20 points)

#Task 14: Essay Question.

You receive a tip that local emergency response times have gotten significantly worse over the past year. You obtain monthly data on response times for police, fire and ambulance calls.

Write 3-4 sentences explaining:
1. What statistical measures would you calculate to verify this claim? (10 points)
2. What visualizations would help readers understand the trends? (5 points)
3. What additional context or data would you need to make this a complete story? (5 points)

**Write your answer here:**

I would look at both the means and medians of this data to get a general impression and compare the two to see the strength of outliers. The standard deviation will also be helpful, as it will allow me to see how far each response time deviates from the average. A histogram would best show this data, as it can categorize the timings in relation to the mean. Providing the source of the data, monthly response times, would be the supporting context I would use for this story.
# Excellent Response


### Submission Instructions
- Save your work frequently
- Make sure all code blocks run without errors
- Provide clear explanations for your analytical choices
- Submit both your code and written responses
- Load this test to GitHub and submit the link in Elms

Good luck!
