---
title: "HW2 Prince George's County Crime"
---

In this assignment, we will be working with a dataset reporting the number of crimes by month in Prince George's County, Maryland, for every month in 2024. 

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv. You will create a dataframe called `pg_crime` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
pg_crime <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv")

```

### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, are any of the categories more important than others? Why or why not? Do any of them make the job of summarizing the data more difficult? Why or why not?

ANSWER HERE: I feel like the data is largely organized in a pretty accessible way, but there are a few things that I'm confused about. I believe I know what is meant by the term "ACCIDENT" but I feel like it is pretty vague and would require further interrogation. If it is a blanket term for a variety of accident-related crimes, I'd like to know what they are. I think cutting up and categorizing the crimes by method (ex. ASSAUL WEAPON vs. ASSAULT SHOOTING) is helpful.


### Task 4: Calculating Averages

Calculate the total number of incidents per month for each category using `mutate()`, calling the new column `total_incidents`. Then calculate the average number of incidents per month for each category of crime, calling the new column `average_incidents`. Assign the result to a new dataframe called `pg_crime_averages`.

```{r}
mutate(pg_crime, total_incidents = rowSums(across(2:13)))
pg_crime <- mutate(pg_crime, total_incidents = rowSums(across(2:13)))

pg_crime_averages <- mutate(pg_crime, average_incidents = total_incidents/12)
```

### Task 5: Describe the Averages

Examine the new dataframe, `pg_crime_averages`, by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the average number of incidents per month for each category. Which categories have the highest average number of incidents per month? Which have the lowest? What does this tell you about the pattern of crime in Prince George's County? Are there categories that you have questions about? Why or why not?

ANSWER HERE:The category with the highest average number of incidents is "accident" and the lowest is "homicide." It seems like the many of the crimes with the highest categories are related to theft or robbery, many of which include cars. I can conclude that residents are most vulnerable to these types of crimes, as they happen the most on average.


### Task 6: Testing a Theory

One theory about crime is that it is more likely to occur in the summer months. Test this theory by creating three new columns using `mutate()` that calculate the percentage change compared to the average for each category of crime for the months of June, July, and August. Call the new columns `june_change`, `july_change`, and `august_change`. Assign the result to a new dataframe called `summer_crime_averages`.

```{r}
mutate(pg_crime_averages, june_change = (7 - average_incidents)/average_incidents * 100, july_change = (8 - average_incidents)/average_incidents * 100, august_change = (9 - average_incidents)/ average_incidents * 100)
```

### Task 7: Describe the Summer Crime Averages

Look at the `summer_crime_averages` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change in the average number of incidents per month for each category of crime for the months of June, July, and August. Do you see any patterns? What does this tell you about the theory that crime is more likely to occur in the summer months?

ANSWER HERE: For the most part, it seems like the theory that crime is more likely to occur in the summer months is incorrect. Most crimes see a decrease in their rates of prevalence besides homicides.

### Task 8: What's the Story?

Write up a lede and nutgraf of a story focused on the crime data in Prince George's County. It could be from any of the dataframes you've created in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

ANSWER HERE Recent analyses of crime data reveal that College Park residents are most vulnerable to auto related crimes, particularly car accidents and motor vehicle theft.

This is based on analysis of recent crime data over the past year and divided by months. The crimes with the highest average rate of occurrence per month are car accidents, motor vehicle theft, and theft from a motor vehicle. 

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
