---
title: "HW3 WH Salaries"
---

In this assignment, we will be working with a dataset containing information about the salaries of White House staffers as of July 2024.

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/biden_wh_salaries_2024.csv. You will create a dataframe called `wh_salaries` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
wh_salaries <- read.csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/biden_wh_salaries_2024.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, what stands out to you about this data? Are there any salary values that will cause problems for analysis?

I think that the sampe size for the data is good, as it helps insulate measurements like the mean from being too skewed based on outliers. With this being said, there are over 10 "0's" in the data, which will still affect the data. 

### Task 4: Calculate the Mean and Median of the Salaries

Write code to calculate the mean and median of the salaries in the dataset using the `summarize()` function. You don't need to make a new dataframe, just display the results.

```{r}
class(wh_salaries$salary)
wh_salaries %>%
summarise(mean_salary = mean(salary, na.rm = TRUE),
median_salary = median(salary, na.rm = TRUE))
```

### Task 5: Add in the minimum and maximum salaries

Take the code you just wrote and add in the minimum and maximum salaries (the functions for that are `min()` and `max()`). Your results should show four columns: mean_salary, median_salary, min_salary, max_salary.

```{r}
wh_salaries %>%
summarise(mean_salary = mean(salary, na.rm = TRUE),
median_salary = median(salary, na.rm = TRUE),
min_salary = min(salary, na.rm = TRUE),
max_salary = max(salary, na.rm = TRUE))
```

### Task 6: Describe the Results

Write a couple of sentences describing the results of your analysis. What do the mean and median salaries tell you about the data? What do the minimum and maximum salaries tell you? Are there any outliers in the data?

I am a bit surprised that the mean and median are as close to eachother as they are. The sample size is a large contributing factor in my opinion, but it still does surprised me that they only vary that little. I am curious about the jobs of those who have a salary of 0, which are outliers, especially because some of them are listed as "directors". I am also curious about what the term "associate counsel" means in this context, and why it demands such a large salary.

### Task 7: Story Idea

Based on the data, what is your best story idea? What is the next step you would take to develop that story idea?

I think the best angle for a story would be looking into why Michelle Barrans's salary is around $60,000 more than the next highest salary. Looking into this story would involve investigating Barrans's role in the White House as well as finding out what is meant by "Associate Counsel." I would then compare her roles to those of others who have the title of "Associate Counsel" and try and determine why she is valued so much more than them.
