---
title: "HW6_Spotify_Songs"
---

In this assignment, you will be working with a dataset of songs played on Spotify in 2020, and you'll be writing code to explore the range of danceability scores in the data, using standard deviation and histograms. You'll also write up your findings.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```

### Task 2: Load the Spotify Songs dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/spotify_2020.csv

```{r}
spotify_songs <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/spotify_2020.csv")
```

### Task 3: Calculate the mean and standard deviation of the danceability scores

Write code to calculate the mean and standard deviation of the danceability scores in the Spotify dataset. Display those.

```{r}
spotify_songs |> summarise(mean_danceability = mean(danceability), sd_danceability = sd(danceability))
```

### Task 4: Create a histogram of the danceability scores

Write code to create a histogram of the danceability scores in the Spotify dataset, based on the example we did in class.

```{r}
   ggplot(spotify_songs, aes(danceability)) +
  geom_histogram()
```

### Task 5: Add a line to the histogram for the mean danceability score

Take the same code and add a vertical line to the histogram for the mean danceability score, based on the example in class.

```{r}
   ggplot(spotify_songs, aes(danceability)) +
    geom_histogram() +
    geom_vline(aes(xintercept = mean(danceability)), color = "red", linetype = "dashed", size = 1)
```


### Task 6: Add two lines to the histogram to show one standard deviation

Take the previous code and add two lines to the histogram for the 1st standard deviation below and above the mean score.

```{r}
ggplot(spotify_songs, aes(danceability)) +
    geom_histogram() +
    geom_vline(aes(xintercept = mean(danceability)), color = "red", linetype = "dashed", size = 1) +
    geom_vline(aes(xintercept = mean(danceability) + sd(danceability)), color = "blue", linetype = "dashed", size = 1) +
   geom_vline(aes(xintercept = mean(danceability) - sd(danceability)), color = "blue", linetype = "dashed", size = 1)
```

### Task 7: Describe our results and pitch a story

Write a few sentences about the distribution of danceability scores in the Spotify dataset. Is it a normal distribution, or is it skewed in one direction? What does the mean and one standard distribution tell us about the songs in the dataset? Draft a short story pitch from this data.

The distribution of the data is not normal, as it is skewed to the left. This means that a majority of the data is on the right side, or more than, the mean. The mean's location coupled with the data within 1 standard deviation of it tells me that most 2/3 of the songs have a danceability ranking of between about .5 and about 86.5. This is a pretty high danceability for most of the data, so I think the best story idea would focus on this how high this is. I could look for commonalities between the most danceable songs in terms of genre or song structure and draw conclusions based on what makes a song "danceable." I can also look for patterns in the artists who make the msot danceable music. A headline I can think of would be something along the lines of "What Makes a Song Danceable? Breaking the Data Down By Trends, Artists and Genres"



### Task 8: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
