---
title: "HW9_TestScores"
Name: Derek Willis/Rob Wells
Date: 10/21/2025
---

name: Luke Rowe


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Did a New Reading Program Lead to Better Scores?

The superintendent recently claimed that a new reading program has improved third-grade reading scores across the school district.

Before the program, third-grade students in the district averaged 72.6 points on standardized reading tests with a standard deviation of 4.8 points.

After implementing the program for one semester, you collected scores from 12 randomly selected classrooms:
74, 76, 73, 75, 78, 77, 74, 79, 75, 76, 77, 75

As a journalist, you need to determine: **Is there statistical evidence that reading scores have actually improved?**

## Task 1: Organize your data and initial assessment

Before you can run this codeblock, you will need to fill in a value where it says REPLACE_ME. That value can be found in the introduction.

```{r}
# Known information about reading scores before the new program
prior_mean <- 72.6  # average score
prior_sd <- 4.8     # standard deviation

# Reading scores after implementing the new program (12 classrooms)
new_scores <- c(74, 76, 73, 75, 78, 77, 74, 79, 75, 76, 77, 75) # Replace with the actual scores

# Create a journalist-friendly dataset
score_data <- tibble(
  classroom = paste("Classroom", 1:12),
  reading_score = new_scores
)

# View the data
score_data
```

### Reflection Question 1:  (2 points)
Based on just looking at the score_data dataframe, have test scores improved? How can you tell?

Tests have improved because every single value is above the mean, which signifies that the new mean would be greater if calculated. 

## Task 2: Calculate key statistics  (2 points)

Like Task 1, you will need to replace values where it says REPLACE_ME before running any code.


```{r}
# Calculate statistics based on the new reading scores
new_stats <- score_data |> 
  summarise(
    mean = mean(new_scores),
    sd = sd(new_scores),
    n = n()
  )

new_stats
```

### Reflection Question 2: (3 points)
Looking at the mean and standard deviation of the new scores compared to the previous statistics, what initial observations can you make? Provide examples from the data in your answer. What questions might these statistics raise for your reporting?

Initially I can see that the mean is is greater than the previous one, but only by 3.15 points. The new mean is 75.75, which could be unimpressive or a sign of good things to come based on how long this "new program" is in place. The standard deviation is less than the previous one, meaning that over 2/3 of the data falls within a range of 1.76455 greater than or less than this mean. The first and most important question I would ask is: "how long has this program been in place for?" This would help me understand just how effective the program has been. Another question I would ask is "what were the highest and lowest scores for these standarized tests?"

## Task 3: Create a column chart  (1 points)

As before, replace any values marked REPLACE_ME based the instructions.


```{r}
# STUDENT TASK: Choose an appropriate fill color for the bars
my_fill_color <- "turquoise" # Replace with a color name like "royalblue", "darkgreen", etc.

# Create a visualization comparing new scores to the previous average
score_data |> 
ggplot(aes(x = classroom, y = reading_score)) +
  geom_col(fill = my_fill_color, alpha = 0.8) +
  geom_hline(yintercept = prior_mean, color = "darkred", size = 1, linetype = "dashed") +
  annotate("text", x = 2, y = prior_mean - 1, 
           label = "Previous Average (72.6)",  hjust = -0.5, vjust= 2, fontface = "bold", color = "darkred") +
  labs(
    title = "Reading Scores After New Program Implementation",
    subtitle = "Horizontal line shows previous district average of 72.6 points",
    x = NULL,
    y = "Reading Test Score",
    caption = "Source: District Assessment Data"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 10),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

### Reflection Question 3: (3 points)
Examine the chart you created, and suggest a better title based on the results of the data, not a description.
"New Reading Program Shows Little-to-No Increase in Reading Scores for Third Graders"

## Task 4: Perform a hypothesis test (2 points)

This is where we formally test the superintendent's claim that reading scores have improved. Fill in the REPLACE_ME values as needed, beginning with your hypotheses.

**Hypotheses:**
Null: The new reading program did not improve reading scores
Alternative: The new reading program improve reading scores

```{r}
# Set the significance level for your test
alpha_level <- .05 # Replace with the appropriate value

# Perform a one-sample t-test
# Since we want to know if scores improved (increased), we use a one-sided test (alternative = "greater")
t_test_result <- t.test(
  score_data$reading_score,
  mu = prior_mean,
  alternative = "greater"
)

# Display the results
t_test_result
```

### Reflection Question 4:  (3 points)
What does the p-value tell you, and what doesn't it tell you? How would you explain these results to a non-technical audience while maintaining accuracy?

The p-value is greater than the significance level of .05, which means that the null hypothesis cannot be rejected. This suggests that the program was ineffective. When explaining this to a non-technical audience, I would say something like "statistical analysis of 12 random scores, each from a different classroom, determined the likelihood that the program  had no effect was significant. This means that there is a notable chance that the program was largely ineffective."

## Task 5: Interpreting the results for your news story (2 points)

Let's gather all of the important stats we'll need in one place, so we can look at the prior average, the new scores and the results of the t.test, including the confidence interval. Replace any values where it says REPLACE_ME.


```{r}
# Get the p-value
p_value <- t_test_result$p.value

# Calculate the 95% confidence interval
ci <- t.test(score_data$reading_score)$conf.int

# Create a tibble to display the key statistics for your story
story_stats <- tibble(
  `Previous average` = prior_mean,
  `New average` = mean(new_scores),
  `Improvement` = mean(new_scores) - prior_mean,
  `Percent change` = round(((mean(new_scores) - prior_mean) / prior_mean) * 100, 1),
  `p-value` = p_value,
  `Lower bound` = ci[1],
  `Upper bound` = ci[2],
  `Confidence level` = "95%"
)

# Display the key statistics
story_stats
```

## Conclusion

### Reflection Question 5: (4 points)
Based on these statistics, what would be your headline and lead paragraph for this story? Is there evidence to support the superintendent's claim?
"Superintendent claims new reading program has increased third great reading scores, the truth is more complicated"

The superintendent recently claimed that the implementation fo a new reading program has increased third-grade reading scores district-wide, but the data reveals a less clear picture. It is true that the average of random samples of 12 test scores, each from a different class, show an increase from the previous average score of around three points. With this being said, statistical analysis of this data shows that there is a significant chance that the program was largely ineffective in this increase.

### Reflection Question 6:  (1 point)
What metrics or outcomes beyond test scores might be important to track for assessing reading performance?
I thinkk that general reading level data could be extremely helpful, as well as data that explores how frequently students read. Another piece of information that I think could help is an anlysis of the curriculum itself and what books the district is using to determine reading levels.
