---
title: "test"
output: html_document
date: "2025-11-12"
---
Name and assignment and date

  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(lubridate)  
library(knitr)     
library(scales) 

```

```{r}
Songs <- read_csv("Billboard_Data.csv") |> 
#rsw comment
  clean_names()
```
Explain divisiveness as the disagreement among three judges. 
Get a mean, median and sd of the rthree judge's ratings.



**The data set I have selected is the Billboard Hot 100 Number Ones Database, which features all the songs that topped the Billboard Hot 100 charts from August 4, 1958 to January 11, 2025. The data was originally compiled by artist and writer Chris Dalla Riva for his book; "Uncharted Territory: What Numbers Tell Us about the Biggest Hit Songs and Ourselves." The database is also instrumental in his weekly newsletter, "Can't Get Much Higher," that explores the intersection between music and data. It can also be found in the Princeton University Library.

Can't Get Much Higher: https://www.cantgetmuchhigher.com/
Princeton University Library: https://dss.princeton.edu/catalog/resource8613

**I think that the most important categories for me will be the "Divisiveness," "Discogs Genre," "Discogs Style" and the "Weeks at Number One" categories. I think that I want to investigate the "Divisiveness" category and its impact on how long songs stay at the top of the Billboard Hot 100. 

**I think that I am missing some context on what exactly "Divisiveness" means for this database, as well as how it was measured. I think that if I can find this out, it will amplify my understanding of the category and inform the methodology of my data analysis. Another thing that could be helpful to know is if artists belonged to races other than black. If I am not mistaken, the database only takes note of if the artist(s) was black, and I think data on the prevalence of other races could be helpful.

```{r}
Songs |> summarize(mean = mean(Divisiveness), median = median(Divisiveness), minimum = min(Divisiveness), maximum = max(Divisiveness))
```

As I mentioned previously, I want to take a look at how a songs divisiveness affected its performance on Hot 100. This requires a thorough analysis of the "Divisiveness," category, and how a song's divisiveness is determined in the first place. For this, I'll have to look into the original dataset as well as Dalla Riva's newsletter to see if I can get a good grasp on what informed the category. Once this is understood I want to use different genres to better understand this category, and see if I can draw correlations between different genres, how divisive they are and if this impacts how long they stay at the top. I think that historical analysis of the music industry and the popular genres at any given time can be helpful for this as well, as it will provide context for how popular the song's genre was at the time it went number one. Finally, I will have to test these correlations for significance using tools like the t-test and ANOVA. This will tell me whether or not I can draw conclusions based on these correlations, and will ultimately inform the structure the story.


NULL HYPOTHESIS: The "Divisiveness" category has no impact on how long a song remained at number 1 on the Billboard Hot 100
ALTERNATE HYPOTHESIS: The "Divisiveness" category has a statistically significant impact on how long a song remained at number 1 on the Billboard Hot 100.

I chose to calculate some advanced statistics for this dataset to better visualize and understand the data. I calculated the margin of error, the mean "Weeks at Number One," the standard error, and both confidence intervals. These stats gave me a fuller set of the data and helped me decide which tests to try and run.

```{r}
divisiveness_summary <- Songs |> 
  summarize(
    mean_divisiveness = mean(Divisiveness),
    sd_divisiveness = sd(Divisiveness),
    mean_weeks = mean(`Weeks at Number One`),
    se_divisiveness = sd_divisiveness / sqrt(1177),
    margin_error = 1.96 * se_divisiveness,
    ci_lower = mean_divisiveness - margin_error,
    ci_upper = mean_divisiveness + margin_error
  )


divisiveness_summary

```

#rsw comment: interpret the results here - what does a 1.55 average mean considering a 0 minimum and a 5 maximum? what does the SD result mean


I chose to use a z-test to find out how each divisiveness score related to the mean. This told me more about variation within the divisiveness category, and how reliable the mean will be. The most the lowest songs all had a Z-score of -1.61, which suggests that they were much lower than the mean. Additionally, the song with the highest Z-score had one of 3.23, which suggest and extremely high score compared to the mean.

```{r}
Divisive_Songs <- Songs |>
  mutate(
    mean_divisiveness = mean(divisiveness),
    sd_divisiveness = sd(divisiveness),
    z_score = (divisiveness - mean_divisiveness) / sd_divisiveness
  ) |> 
  select(song, artist, date, weeks_at_number_one, divisiveness, z_score,  mean_divisiveness,  sd_divisiveness)
```

#rsw comment: what does it tell you about the hypothesis

What does divisiveness mean and how is it presented in the data
NULL Hypothesis: There is little to no variation in divisivness
ALTERNATE HYPOTHESIS: XXXX



code failed
```{r}

t.test(weeks_at_number_one ~ divisiveness, var.equal=TRUE, data = Divisive_Songs)
```
I wanted to perform a t test as well to test the relationship between the variables, but had an extremely difficult time getting the code to work. I spent all day using the activities (car_thefts, Margin of Error - Housing Prices) to help me try and run this t-test correctly, but I haven't been able to figure this out. I initially had errors about the pipe I was using , which through research I was able to fix but then R was having issues locating the "Divisiveness" category. I genuinely do not know why this is, as I tried every method we used in class to identify the dataset and relevant variables to no avail. I thought I had a decent grasp on the t-test, but this experience has shown me I still need to do more to be able to replicate them with this dataset.



original
```{r}

Relationship <- Songs |>
  ggplot(mapping =aes(x=`Weeks at Number One`, y=Divisiveness)+ 
           geom_boxplot()+
           geom_jitter(width=2))
  show(Relationship)
```


```{r}
Songs |>
  ggplot(aes(x=weeks_at_number_one, y=divisiveness)) + 
         geom_boxplot() +
         geom_jitter(width=2)

```
I tried to create a box plot but after hours of working with the variables I have been unable to make anything show up. I think the problem has to do with a syntax error, but I am genuinely confused as to why nothing is coming up. I tried to consult the activities and homework and use them as a model for how I made this boxplot, but I continued to run into errors. I also consulted my brother who uses R for his job, as well as another friend who has experience with the program and each of us were unable to figure it out. I think it has to do with the nature of the variables and the size of the dataset, so I think I need to find a way to only use a sample of it for these plots to work.
